---
title: "Shiny in production"
subtitle: "Brighton R"
author: "Chris Beeley"
institute: "Nottinghamshire Healthcare NHS Foundation Trust"
date: "8th April 2021"
output:
  xaringan::moon_reader:
    css: 
      - default
      - css/nhsr.css
      - css/nhsr-fonts.css
    lib_dir: libs
    seal: false
    self_contained: true
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
---
class: title-slide, left, bottom

```{r setup, include = FALSE}

library(knitr)
library(tidyverse)
library(nhsrtheme)
library(xaringanExtra)
library(emojifont)

# set default options

opts_chunk$set(echo = FALSE,
               fig.width = 7.252,
               fig.height = 4,
               dpi = 300,
               dev.args = list(type = "cairo"),
               echo = TRUE,
               warning = FALSE,
               message = FALSE,
               error = FALSE)

# uncomment the following lines if you want to use the NHS-R theme colours by default
scale_fill_continuous <- partial(scale_fill_nhs, discrete = FALSE)
scale_fill_discrete <- partial(scale_fill_nhs, discrete = TRUE)
scale_colour_continuous <- partial(scale_colour_nhs, discrete = FALSE)
scale_colour_discrete <- partial(scale_colour_nhs, discrete = TRUE)

library(xaringanExtra)
xaringanExtra::use_share_again() # need to get the slide button on html view

```

```{r echo=FALSE}

xaringanExtra::use_logo(
  image_url = "img/nottshc.png",
  exclude_class = c("title-slide", "inverse", "hide-logo"),
  width = "300px",
  height = "200px",
  position = css_position(top = "0em", right = "1em")
)

```


# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author` | `r rmarkdown::metadata$date`
</br>

---

# Introduction

* This presentation is about Shiny in production
* It’s deliberately pitched at people who are using Shiny already and want to use it better
* If you’re not a Shiny user yet hopefully you will still get an understanding of developments in Shiny

---

# The bad old days

* I started using Shiny very early on
* Most of the development at the time was based on adding functionality and increasing the power of Shiny
* It was a time of rapid development but it led to:
* Enormous (2000+ line) server.R files
* Difficult to debug
* Difficult to test
* Difficult to collaborate on
* Difficult to deploy

---

# The new wave

* NONE of these ideas are mine
* Big credit goes to Colin Fay and the team at [ThinkR](https://thinkr.fr/) 
* In particular [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/) 
* And [Mastering Shiny](https://mastering-shiny.org/)
* I’m going to be opinionated with other people’s opinions `r emoji("wink")`
* (My team and I are not perfect- but we’re trying to be `r emoji("wink")`)

---

# Functions!

* Everybody knows you should use functions…
    * Don’t they?
* Ideally a Shiny application should consist ONLY of reactive code
* There should be NO computation (“business logic”) in server.R
* Even if you only do it once, put it in a function
* Functions are self documenting
* Functions are easy to test
* More recently Shiny has added automatic running of code in R/ folder (1.5?)

---

# Packages

* Documented
* Tested
* Shareable
* Business logic in server.R `r icons::fontawesome("arrow-right")` business logic in functions `r icons::fontawesome("arrow-right")` business logic in team package

---

# Modules

* Modules allow you to reuse Shiny code

<iframe src="https://involve.nottshc.nhs.uk/rsconnect/experience_a/" width="1000" height="400"></iframe>

---

# But wait, there's more

* Modules are _namespaced_
* This makes it easier to collaborate with *yourself*
* The unit of collaboration becomes the module
* I've never collaborated on a Shiny app without modules and I never want to

---

# Scoping data

* This is a blessing and a curse

```{r, eval=FALSE, echo=TRUE}

endo_data <- mod_clean_and_merge_server("clean_and_merge_ui_1", 
                                        header_filename = "endo.rda")
  
path_data <- mod_clean_and_merge_server("clean_and_merge_ui_2", 
                                        header_filename = "path.rda")

# in module

reactive(
  endoData()
)

```

---

# Passing data around

```{r, eval=FALSE, echo=TRUE}

# server.R

merge_data <- mod_merge_data_server("merge_data_ui_1", endo_data = endo_data)

# module

mod_merge_data_server <- function(id, endo_data, path_data, load_prev){
  moduleServer( id, function(input, output, session){
    # ...

    selectInput(session$ns("endoDate"), "Endoscopy date", 
                           choices = names(endo_data()))
}

```

---

# Gitflow

* This isn't Shiny specific but we try to adopt a clear approach to gitflow and GitHub
* Issues are filed for each update/ bugfix and given a branch with the issue number
* Pull requests are made against the development branch and merged by the "lead" for each repo
* We use semantic versioning for all packages and apps and branch dev to a release branch
* The master is "clean" and should always work
* Using namespaced modules helps us to collaborate on GitHub easily

---

# {golem}

> [{golem}](https://github.com/ThinkR-open/golem) is an opinionated framework for building production-grade shiny applications

* In golem, all Shiny applications are packages
* Packages make it easy to test, manage dependencies, and deploy

```{r, eval=FALSE, echo=TRUE}
# install.packages("remotes")
remotes::install_github("CDU-data-science-team/cmhsDashboard")
library(cmhsDashboard)
run_app()

```

* This is magic!

---

# Golem is your new best friend

* Golem includes lots of code to make writing Shiny applications and packages easier
* {usethis}
    * Licence
    * Dependencies
    * data/ folder
    * Git
* {testthat}
* {golem}
    * Favicons
    * Add modules
    * Deploy RStudio Connect/ Shiny server/ shinyapps/ Docker

---

# Package dependencies

* Managed more easily because of packaging
* Golem encourages, but doesn't mandate, making explicit function calls from packages

```{r, eval = FALSE, echo = TRUE}

df <- data %>%  
  tidyr::unnest(question_type) %>% 
  dplyr::mutate(dplyr::across(all_of(question_type), as.numeric)) %>% 
  dplyr::left_join(category_table, join_lookup)

```

* This is again extremely useful for collaborating

---

# Golem YAML

* It's early days for us with this but this is a very powerful and simple tool

```{r, eval=FALSE, echo=TRUE}

default:
  golem_name: staffSurveyNew
  golem_version: 0.0.0.9000
  app_prod: no
  where: github #<<
production:
  app_prod: yes
dev:
  golem_wd: !expr here::here()

```

---

# In the app

```{r, eval=FALSE, echo=TRUE}

where_am_i <- get_golem_config("where")

if(where_am_i == "github"){
  
  # do something
} else {
  
  # do something else
}


```


---

class: inverse

# Contact

[`r icons::fontawesome("twitter")` @ChrisBeeley](https://twitter.com/ChrisBeeley)<br/>
[`r icons::fontawesome("github")` Clinical Development Unit Data Science Team](https://github.com/CDU-data-science-team)<br/>
[`r icons::fontawesome("paper-plane")` chris.beeley@notthshc.nhs.uk](mailto:chris.beeley@nottshc.nhs.uk)

